---
title: "STAT495 (Advanced Data Analysis)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "November 7, 2016"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
library(mdsr)
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
library(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

### SQL and dplyr

```{r message=FALSE}
library(RMySQL)
db <- src_scidb("airlines")
flights <- tbl(db, "flights")
carriers <- tbl(db, "carriers")
```

```{r warning=FALSE}
q <- flights %>%
  filter(year == 1996 & month == 9 & dest == "JFK") %>%
  inner_join(carriers, by = c("carrier" = "carrier")) %>%
  group_by(carrier) %>%
  summarize(N = n(), pct_ontime = sum(arr_delay <= 15) / n()) %>%
  filter(N >= 100) %>% arrange(desc(pct_ontime))
head(q, 3)
```

Note that we can get (somewhat ugly) SQL translations from dplyr.

```{r}
show_query(q)
```

```{r}
query <- "SELECT *
FROM (SELECT *
FROM (SELECT `carrier`, count(*) AS `N`, SUM(`arr_delay` <= 15.0) / count(*) AS `pct_ontime`
FROM (SELECT * FROM (SELECT *
FROM (SELECT *
FROM `flights`
WHERE (`year` = 1996.0 AND `month` = 9.0)) `gzmlvcdimf`
WHERE (`dest` = 'JFK')) `sxzfnmbiww`
INNER JOIN `carriers` USING (`carrier`)) `axvcucedqr`
GROUP BY `carrier`) `wxtkiyemoe`
WHERE (`N` >= 100.0)) `qrkfikumzk`
ORDER BY `pct_ontime` DESC"
```

```{r warning=FALSE}
ds <- DBI::dbGetQuery(db$con, query); head(ds, 3)
```

In groups of three, please work through the following exercises.
Each of the following exercises can be solved via a single SQL query. Equivalently, each can be solved via a single pipeline of `dplyr` commands. Write valid solutions using both methods (please don't use `show_query()` to generate the SQL).


#### Group Blue:
How many domestic flights flew into Dallas-Fort Worth (`DFW`) on May 14, 1998?

SOLUTION 1 (dplyr):

```{r}

```

SOLUTION 2 (SQL):

```{r}


```

#### Group Green: 
Which airport had the highest average arrival delay time in 2008?

SOLUTION 1 (dplyr):

```{r}

```

SOLUTION 2 (SQL):

```{r}


```

#### Group Orange: 
Of all the destinations from Chicago O'Hare (`ORD`), which were the most common in 1997?

SOLUTION 1 (dplyr):

```{r}
NICK <- flights %>%
  filter(year == 1997, origin == "ORD") %>%
  group_by(dest) %>%
  summarize(counts = n()) %>%
  arrange(desc(counts)) 

NICK <- data_frame(NICK)
head(NICK)
```

SOLUTION 2 (SQL):

```{r}
queryChicago <- "
SELECT origin, dest, year, count(*) AS counts
FROM flights 
WHERE `origin` = 'ORD' AND `year` = 1997.0
GROUP BY `dest`
ORDER BY `counts` DESC"


ds <- DBI::dbGetQuery(db$con, queryChicago)
head(ds, 3)

queryPUB <- "SELECT * FROM airports WHERE faa = 'PUB'"
dsPUB <- DBI::dbGetQuery(db$con, queryPUB)
dsPUB
```

#### Group Purple:
Find all flights between `JFK` and `SFO` in 1994. How many were cancelled? What percentage of the total number of flights were cancelled?

SOLUTION 1 (dplyr):

```{r}

```

SOLUTION 2 (SQL):

```{r}


```
